---
title: "API practice with R"
output: html_document
date: "2022-08-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dataRetrieval)
library(tidyverse)
library(janitor)
library(metajam)
```

## Part 1: Plotting Ventura River discharge from USGS data

Use the USGS `dataRetrieval` R package to retrieve hydrological data.

Tutorial on: https://cran.r-project.org/web/packages/dataRetrieval/vignettes/dataRetrieval.html#daily-data

Plot the discharge time-series for the Ventura River from 2019-10-01 to 2020-10-05

```{r}
#Pull Ventura river data - site 11118500

siteNumber <- "11118500"
ChoptankInfo <- readNWISsite(siteNumber)
parameterCD <- "00060"

#Raw daily data

rawDailyData <- readNWISdv(siteNumber, parameterCD, "2019-10-01", "2020-10-05")


#Look at the metadata to make sense of this:
siteInfo <- attr(rawDailyData, "siteInfo")
#that was unhelpful...

#rename columns 
ventura_discharge <- renameNWISColumns(rawDailyData) %>% 
  clean_names()

names(ventura_discharge)


```

Now it's time to plot:

```{r}

ggplot(data = ventura_discharge, aes(x = date,
                                     y = flow)) +
  geom_jitter(size = 0.5, aes(color = flow))


```

## Part 2: Using `metajam`

The metajam R package relies on the dataONE API to download data and metadata into your R Environment. It is currently supporting KNB, ADC and EDI repositories because they rely on the metadata standard EML.

Intro to metajam: https://brunj7.github.io/EDS-214-analytical-workflows/slides/metajam_intro.html#1

### determine what percentage of Alaskan households are only speaking English:

```{r}
#set inputs

data_obj <- "https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3A7fc6f6db-c5ea-426a-a743-1f2edafb43b8"
path <- "~/eds214-day3-api-notes"

#downloaddata and metadata

download_d1_data(data_obj, path)

#now read in the downloaded files - this returns a list of dataframes

full_dataset <- read_d1_files("~/eds214-day3-api-notes/doi_10.5063_F1CJ8BPH__household_language__csv")

#Extract the dataframe

language_data <- full_dataset$data %>% 
  clean_names()


```

Now write code that will compute the percentage of Alaskan households speaking only english for 2009 to 2015

```{r}
data_subset <- language_data %>% 
  filter(year %in% 2009:2015) %>% 
  group_by(year) %>% 
  summarize("only-english-percentage" = (sum(speak_only_english, na.rm = TRUE)/sum(total, na.rm = TRUE))*100) #%>% 
  #mutate(year = lubridate::year(year))

plot_data <- as.data.frame(data_subset)

```

and plot to visualize

```{r}

ggplot(data = plot_data, aes(x = "year",
                             y = "only-english-percentage")) +
         geom_jitter() +
  #xlim(2009, 2015) +
  #ylim(0, 100)

  
```









